A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-09 paper/paper1/ms.tex
Revising the introduction

Atmospheric escape appears to be a compulsory phase for exoplanets of a certain size and insolation, statistically imprinted in the dearth of 1.5-2.0~$R_\oplus$ planets \citep{2017AJ....154..109F}.  The two physical mechanisms for this rapid transition between mini-Neptunes and super-Earths include photoevaporation \citep{2013ApJ...775..105O,2017ApJ...847...29O} and core-powered mass loss \citep{2019MNRAS.487...24G}.  Whatever the cause, some large fraction of planets must undergo atmospheric escape and the signal should be widely discernable, at least in observationally favorable configurations such as transiting planets.

Enough uncertainty exists in the tradeoffs among physical mechanisms, system ages, and timescales to make it unclear whether any given exoplanet will be actively undergoing atmospheric escape.  The most favorable traits may include close-in, low surface gravity exoplanets within a strategically chosen radius range.  The 1.5-2.0 Earth radius range could be expected to harbor atmospheres that have not yet shed their gaseous envelopes, still in the act of descending to their final state.  Alternatively, inflated Hot Jupiters could be so dramatically ablated as to overcome the somewhat stronger potential well.  Significant gains in understanding can therefore be expected from observational studies sampling the extremes of the exoplanet distribution.

Candidates for detection should be observable in one of three atomic lines.  Ly~$\alpha$, \ion{He}{1} 10833 \AA, and H~$\alpha$ have emerged as the most amenable to detection \citep{2018ApJ...855L..11O,2023MNRAS.518.4357O}.  Ly $\alpha$ can exhibit strong atmospheric escape signals, but instellar medium \ion{H}{1} absorption censors its low-velocity linecore.  \ion{He}{1} 10833 \AA~ offers significant advantages over Ly~$\alpha$, and has resulted in 8 systems with detections to date: \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N, 2021A&A...647A.129L}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A,2020AJ....159..115K}, HD 209458 b \citep{2019A&A...629A.110A}, WASP 69 b \citep{2020AJ....159..278V}, WASP 52 b \citep{2020AJ....159..278V}, HD 189733 b \citep{2021A&A...647A.129L}, and HAT-P-18 b \citep{2021ApJ...909L..10P}.  These detections exhibit a range of signal strengths, as high as 10$\%$ transit depths.  Measurements at high spectral resolution have been especially valuable for evincing velocity substructure of the escaping gas motion relative to the exoplanet restframe \citep{2019A&A...629A.110A,2020ApJ...894...97N}, and convincingly associating the signal to an exoplanetary origin as opposed to stellar contamination \citep{2018AJ....156..189C}.

>>>>

Atmospheric escape appears to be a compulsory phase for exoplanets of a certain size and insolation, statistically imprinted in the dearth of 1.5-2.0~$R_\oplus$ planets \citep{2017AJ....154..109F}.  The leading candidate physical mechanisms for this rapid transition between mini-Neptunes and super-Earths include photoevaporation \citep{2013ApJ...775..105O,2017ApJ...847...29O} and core-powered mass loss \citep{2019MNRAS.487...24G}.  Whatever the cause, some large fraction of planets must undergo atmospheric escape and the signal should be widely discernable, at least in observationally favorable configurations such as transiting planets.  At least 28 examples have been seen to date \citep{2022arXiv221116243D}.

Uncertainty in system ages, evaporation timescales, and dominating physical mechanisms, degrade our ability to foretell if any given planet will exhibit ongoing signature of atmospheric escape.  Episodic stellar wind gusts and other forms of cosmic variance could also subdue atmospheric escape, even when we expect it.  The 48 published non-detections of atmospheric escape \citep{2022arXiv221116243D} encode these natural whims in a way that we have not yet disentangled. Nevertheless, we can boost our chances of witnessing active and significant atmospheric escape by tailoring our survey to sources with our notions of favorable traits.  These traits may include close-in, low surface gravity exoplanets within a strategically chosen radius range.  The 1.5-2.0 Earth radius range could be expected to harbor atmospheres that have not yet shed their gaseous envelopes, still in the act of eroding to their final rocky state.  Alternatively, inflated Hot Jupiters could be so dramatically ablated as to overcome the somewhat stronger potential well and facilitating hydrodynamic flows.  Significant gains in understanding can therefore be expected from observational studies sampling the extremes of the exoplanet distributions.

Candidates for detection should usually be observable in one of three atomic lines.  Ly~$\alpha$, \ion{He}{1} 10833 \AA, and H~$\alpha$ have emerged as the most amenable to detection \citep{2018ApJ...855L..11O,2023MNRAS.518.4357O}.  Ly $\alpha$ can exhibit strong atmospheric escape signals, but instellar medium \ion{H}{1} absorption censors its low-velocity linecore.  \ion{He}{1} 10833 \AA~ offers significant advantages over Ly~$\alpha$, and has resulted in at least 14 systems with detections to date, including \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N, 2021A&A...647A.129L}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A,2020AJ....159..115K}, HD 209458 b \citep{2019A&A...629A.110A}, WASP 69 b \citep{2020AJ....159..278V}, WASP 52 b \citep{2020AJ....159..278V}, HD 189733 b \citep{2021A&A...647A.129L}, and HAT-P-18 b \citep{2021ApJ...909L..10P}.  These detections exhibit a range of signal strengths, as high as 10$\%$ transit depths.  Measurements at high spectral resolution have been especially valuable for evincing velocity substructure of the escaping gas motion relative to the exoplanet restframe \citep{2019A&A...629A.110A,2020ApJ...894...97N}, and convincingly associating the signal to an exoplanetary origin as opposed to stellar contamination \citep{2018AJ....156..189C}.
